---
title: "Data imputation"
author: "Marta"
date: "April 9, 2019"
output:
  pdf_document: default
  html_document: default
---
Clean up global variables (used for debuggining)
```{r}
rm(list=ls())
```


Package used for the stochatic regression imputtion is called mice. (takes a bit of time but needed to do that every time)
```{r  include=FALSE}
install.packages("mice")

```

Reading in the data. Don't know how to do all at once so I'll do that later

```{r}
stats <- read.csv("01_compressed.csv", header=TRUE, sep=",")
stats[12] <- NULL #drop games
stats[12] <- NULL #drop games
stats[18] <- NULL #drop weather
#head(stats)
```

Check data for missing values

```{r}
sapply(stats, function(x) sum(is.na(x)))
```

Setting each vatiable. (need to install that package here. Might not work the first time you run, just run again, its the installation problem)

```{r}
install.packages("dplyr")
library(dplyr) 
stats <- stats %>%
    mutate(
        time = as.Date(time),
        mood = as.numeric(mood),
        circumplex.arousal = as.numeric( circumplex.arousal),
        circumplex.valence = as.numeric(circumplex.valence),
        activity = as.numeric(activity),
        screen = as.numeric(screen),
        call = as.numeric(call),
        sms = as.numeric(sms),
        appCat.builtin = as.numeric(appCat.builtin),
        appCat.communication = as.numeric(appCat.communication),
        appCat.entertainment = as.numeric(appCat.entertainment),
        #appCat.finance  = as.numeric(appCat.finance ),
        #appCat.game = as.numeric(appCat.game),
        appCat.office = as.numeric(appCat.office),
        appCat.other = as.numeric(appCat.other),
        appCat.social = as.numeric(appCat.social),
        appCat.travel  = as.numeric(appCat.travel ),
        appCat.unknown = as.numeric(appCat.unknown),
        appCat.utilities = as.numeric(appCat.utilities),
        appCat.utilities   = as.numeric(appCat.utilities ),
        #appCat.weather = as.numeric(appCat.weather),
        day_0 = as.integer(day_0),
        day_1 = as.integer(day_1),
        day_2 = as.integer(day_2),
        day_3 = as.integer(day_3),
        day_4 = as.integer(day_4),
        day_5 = as.integer(day_5),
        day_6 = as.integer(day_6)
                                        
    )
```

Just to check if everything is correct

```{r}
str(stats)
```
select the variable to not be included as predictor during imputation.
```{r}
library(mice)

init = mice(stats, maxit=0) 
meth = init$method
predM = init$predictorMatrix
```

Don't take this under consideration while predicting data:

```{r}
predM[, c("time")]=0
predM[, c("day_0")]=0
predM[, c("day_1")]=0
predM[, c("day_2")]=0
predM[, c("day_3")]=0
predM[, c("day_4")]=0
predM[, c("day_5")]=0
predM[, c("day_6")]=0
```

Skipping variable for imputing - YOU MIGHT NEED TO ADD SOMETHING HERE

```{r}
#meth[c("appCat.weather")]=""
#meth[c("appCat.game")]=""
```

Setting method for imputing - choose which method do you want the variables to be predicted with
https://www.rdocumentation.org/packages/mice/versions/3.4.0/topics/mice


```{r} 
meth[c("mood")]="norm.boot" 
meth[c("circumplex.arousal")]="norm.boot" 
meth[c("circumplex.valence")]="norm.boot"
meth[c("activity")]="norm.boot" 
meth[c("screen")]="norm.boot" 
meth[c("call")]="norm.boot"
meth[c("sms")]="norm.boot" 
meth[c("appCat.builtin")]="norm.boot" 
meth[c("appCat.communication")]="norm.boot"
meth[c("appCat.entertainment")]="norm.boot" 
#meth[c("appCat.finance")]="norm.nob" 
meth[c("appCat.office")]="norm.boot" 
meth[c("appCat.other")]="norm.boot" 
meth[c("appCat.social")]="norm.boot"
meth[c("appCat.travel")]="norm.boot" 
meth[c("appCat.unknown")]="norm.boot" 
meth[c("appCat.utilities")]="norm.boot"

```


**This is the imputation part**

```{r}
set.seed(500)
imputed = mice(stats, method='cart', predictorMatrix=predM, m=5)
```

Saving the imputed data

```{r}
imputed <- complete(imputed) 
```

In here I check if everything was filled in 

```{r}
sapply(imputed, function(x) sum(is.na(x)))
```

Write predicted data to a csv

```{r}
write.table(imputed, file = "01_imputed.csv", sep = ", ", row.names=FALSE) #print into csv
```

```{r}
rm(list=ls())

inpath = "01_compressed.csv"
outpath = "01_imputed.csv"

stats <- read.csv(inpath, header=TRUE, sep=",")
#stats[12] <- NULL #drop games
#stats[12] <- NULL #drop games
#stats[18] <- NULL #drop weather

library(dplyr) 
stats <- stats %>%
    mutate(
        time = as.Date(time),
        mood = as.numeric(mood),
        circumplex.arousal = as.numeric( circumplex.arousal),
        circumplex.valence = as.numeric(circumplex.valence),
        activity = as.numeric(activity),
        screen = as.numeric(screen),
        call = as.numeric(call),
        sms = as.numeric(sms),
        appCat.builtin = as.numeric(appCat.builtin),
        appCat.communication = as.numeric(appCat.communication),
        appCat.entertainment = as.numeric(appCat.entertainment),
        appCat.finance  = as.numeric(appCat.finance ),
        appCat.game = as.numeric(appCat.game),
        appCat.office = as.numeric(appCat.office),
        appCat.other = as.numeric(appCat.other),
        appCat.social = as.numeric(appCat.social),
        appCat.travel  = as.numeric(appCat.travel ),
        appCat.unknown = as.numeric(appCat.unknown),
        appCat.utilities = as.numeric(appCat.utilities),
        appCat.utilities   = as.numeric(appCat.utilities ),
        appCat.weather = as.numeric(appCat.weather),
        day_0 = as.integer(day_0),
        day_1 = as.integer(day_1),
        day_2 = as.integer(day_2),
        day_3 = as.integer(day_3),
        day_4 = as.integer(day_4),
        day_5 = as.integer(day_5),
        day_6 = as.integer(day_6)
                                        
    )

library(mice)

init = mice(stats, maxit=0) 
meth = init$method
predM = init$predictorMatrix

predM[, c("time")]=0
predM[, c("day_0")]=0
predM[, c("day_1")]=0
predM[, c("day_2")]=0
predM[, c("day_3")]=0
predM[, c("day_4")]=0
predM[, c("day_5")]=0
predM[, c("day_6")]=0

#meth[c("appCat.weather", "appCat.game")]=""
#meth[c("appCat.game")]=""

meth[c("mood")]="norm.boot" 
meth[c("circumplex.arousal")]="norm.boot" 
meth[c("circumplex.valence")]="norm.boot"
meth[c("activity")]="norm.boot" 
meth[c("screen")]="norm.boot" 
meth[c("call")]="norm.boot"
meth[c("sms")]="norm.boot" 
meth[c("appCat.builtin")]="norm.boot" 
meth[c("appCat.communication")]="norm.boot"
meth[c("appCat.entertainment")]="norm.boot" 
meth[c("appCat.finance")]="norm.nob" 
meth[c("appCat.office")]="norm.boot" 
meth[c("appCat.other")]="norm.boot" 
meth[c("appCat.social")]="norm.boot"
meth[c("appCat.travel")]="norm.boot" 
meth[c("appCat.unknown")]="norm.boot" 
meth[c("appCat.utilities")]="norm.boot"

set.seed(500)
imputed = mice(stats, method='cart', predictorMatrix=predM, m=5)

imputed <- complete(imputed) 
#imputed[is.na(imputed)] <- 0 # substituting NaN values of empty columns for 0
write.table(imputed, file = outpath, sep = ",", row.names=FALSE) #print into csv
```




